### 2015-07-20 shrink/design.tex
added abstract, small reference fixes in rest of thesis.
a finished version: jul 20, 11 AM

\section{\shrink\ design and implementation}
\label{sec:design}

Shrink is a system for server and network consolidation as well as load balancing. 
\shrink\ reduces a \cdc's energy use, while enabling operators to achieve the desired response times and hardware reliability.
A novel aspect of \shrink\ is its network-aware server consolidation scheme, which increases the energy savings that \shrink's network consolidation scheme achieves. This section describes \shrink's design goals, its design and its implementation.

@@

(2) \textbf{Response time:} The design should enable operators to provide the desired response time to end users.

@@

\shrink\ runs as a control program that executes at fixed length intervals. In each interval, the control program receives  reports from all active servers regarding their load at the granularity of \emph{content buckets}; content are mapped to buckets using a consistent hash function based on their name.  
Based on three inputs  -- server load reports, a utilization vs. end-user response time curve for the operator's response time metric of interest (such as a given percentile of response time) and a specified value of end-user response time as per that metric -- , the control program computes the following two outputs: (1) \textbf{Consolidation:} The set of servers and switches to keep active, while remaining servers and switches are turned off. (2) \textbf{Load balancing:} A \emph{traffic split} for each bucket describing the ratio in which the bucket's requests are split among servers. We next describe \shrink's consolidation and load balancing algorithms, followed by how to compute the utilization vs. end-user response time curve provided to \shrink.
%
%
%server load reports, 
%
%
%
%a end-user response time metric specified by the operator (such as a given percentile of response time) and 
%
%

@@

%, followed by how to compute the utilization-vs-response time curve provided to \shrink.

@@

%Choosing the highest server utilization limit ($U$) that still achieves the desired end-user response times maximizes energy savings for operators. Unfortunately, the relation between $U$ and the response time  of \shrink\ is challenging to obtain without prior testing for a representative workload. To appreciate the challenge, consider this simple but  inaccurate way to select $U$. Given a desired response time ($r_0$) and the relation $r = f(u)$, which provides the steady-state response time $r$ of a server at an utilization $u$, choose $U = f^{-1}(r_0)$. The $U$ thus selected will likely result in a response time that is higher than the desired for several reasons such as imperfect load balancing, coordination overhead among peer caches and non-steady state behavior resulting from on-off transitions. The relation between $U$ and the response time of \shrink\ can be obtained with prior testing for a representative workload, thereby enabling operators to choose $U$ that maximizes energy savings while achieving desired response times.

@@

%How does an operator select the utilization limit $U$ that maximizes energy savings while achieving the desired response time? A simple, but somewhat inaccurate, way to select $U$ is as follows. Given a desired response time ($r_0$) and a steady-state utilization-vs.-response time curve of a server $r = f(u)$, choose $U = f^{-1}(r_0)$. The utilization limit selected thus will likely result in a response time that is higher than the desired value due to imperfect load balancing and non-steady state behavior immediately after servers have been turned on or off. However, this utilization limit is expected to be a good starting point, after which operators can reduce $U$ in small decrements until response time reaches the desired value.

@@

Let $F(u)$ be the utilization vs. end-user response time curve for the operator's response time metric of interest provided to \shrink. $F(u)$ is similar to the end-to-end response time $(f(u) + mB + A)$ described in Section \ref{sec:analysis}. We discuss how $F(u)$ can be obtained in Section \ref{sec:utilization-vs-responsetime}.  \shrink\ uses $F(u)$ assuming that if the servers active in the datacenter are run at an average utilization $u$, then the resulting response time is $F(u)$. Thus, to achieve a specified end-user response time $R$, \shrink\ chooses a server utilization limit $U = r^{-1}(R)$.

%\shrink\ chooses a server utilization limit $U$ to achieve a specified end-user response time $R$.  If $F(u)$ is the utilization vs. end-user response time curve provided, $U = r^{-1}(R)$. $F(u)$ is similar to the end-to-end response time $(f(u) + mB + A)$ described in Section \ref{sec:analysis}. We discuss how $F(u)$ can be obtained in Section \ref{sec:utilization-vs-responsetime}.

@@

\subsection{Computing utilization vs. response time curve}
\label{sec:utilization-vs-responsetime}

We discuss three ways of computing the utilization vs. response time curve $F(u)$ input to \shrink\ in the order of increasing complexity. The more complex schemes potentially enable \shrink\ to provide response times that are closer to those specified by operators.

The first and second approaches require prior measurements of a single server and of the entire \cdc\ respectively.
These measurements obtain the  value of the operator's response time metric of interest at varying levels of utilization.  These measurements are done with a representative workload in a test environment with real or emulated client-to-server delay and server-to-origin delay.  The first approach implicitly assumes that if the response time of a single server at utilization $u$ is $r$, then the response time of the \cdc\ whose servers have an average utilization $u$ is also $r$. This assumption ignores the inefficiencies of a distributed system such as imperfect load balancing and coordination overhead among peer caches in a \cdc. These factors are more accurately accounted for in the second approach. Nonetheless, both approaches could be inaccurate as they do not consider the non-steady cache behavior due to on-off transitions and increased cache miss rates due to reduced storage.  To account for these inaccuracies, a constant inefficiency factor $\rho$ is added as follows. If $F'(u)$ is the measured  utilization vs. response time curve, then \shrink\ is input a curve $F(u) = \rho \times F'(u)$.

The third approach is to equip \shrink\ to learn the curve automatically based on online measurements. The overhead of these measurements can be kept small by doing them for a small fraction of requests only. This approach could be more accurate than the previous two approaches because it considers the factors discussed above and can even adapt to changes in workload characteristics that influence the utilization vs. response time behavior. We have not implemented this third approach in \shrink.

%We discuss three ways of computing the utilization vs. response time curve that is provided to \shrink\ in the order of increasing complexity. The more complex schemes potentially enable \shrink\ to provide response times that are closer to those specified by operators. The first and second approaches require prior measurements at varying levels of utilization of a single server and of the entire \cdc\  respectively. Similar to to the assumption made in deriving the model in Section \ref{sec:analysis}, the first approach implicitly assumes that the utilization vs. response time behavior of the datacenter is the same as that of a single server. However, a \cdc\ will likely have a higher response time than a single server at the same utilization because imperfect load balancing and coordination overhead among peer caches in a \cdc.  These factors are accounted for in the second approach; therefore, its utilization vs. response time curve is closer to the actual curve than the first approach.

%The first, and the simplest, approach is to compute the utilization vs. response time curve based on profiling a single server at varying levels of utilization. This approach implicitly assumes that the utilization vs. response time behavior of the datacenter is the same as that of a single server; we make the same assumption in Section \ref{sec:multiserver}.

>>>>

\section{\shrink\ design and implementation}
\label{sec:shrink-design}

Shrink is a system for server and network consolidation as well as load balancing. 
\shrink\ reduces a \cdc's energy use, while enabling operators to achieve the desired latencies and hardware reliability.
A novel aspect of \shrink\ is its network-aware server consolidation scheme, which increases the energy savings that \shrink's network consolidation scheme achieves. This section describes \shrink's design goals, its design and its implementation.

@@

(2) \textbf{User-perceived latency:} The design should enable operators to provide the desired latency to end users.

@@

\shrink\ runs as a control program that executes at fixed length intervals. In each interval, the control program receives  reports from all active servers regarding their load at the granularity of \emph{content buckets}; content are mapped to buckets using a consistent hash function based on their name.  
Based on three inputs  -- server load reports, a utilization vs. end-user latency curve for the operator's latency metric of interest (such as a given percentile of latency) and a specified value of end-user latency as per that metric -- , the control program computes the following two outputs: (1) \textbf{Consolidation:} The set of servers and switches to keep active, while remaining servers and switches are turned off. (2) \textbf{Load balancing:} A \emph{traffic split} for each bucket describing the ratio in which the bucket's requests are split among servers. We next describe \shrink's consolidation and load balancing algorithms, followed by how to compute the utilization vs. end-user latency curve provided to \shrink.
%
%
%server load reports, 
%
%
%
%a end-user latency metric specified by the operator (such as a given percentile of latency) and 
%
%

@@

%, followed by how to compute the utilization-vs-latency curve provided to \shrink.

@@

%Choosing the highest server utilization limit ($U$) that still achieves the desired end-user latencies maximizes energy savings for operators. Unfortunately, the relation between $U$ and the latency  of \shrink\ is challenging to obtain without prior testing for a representative workload. To appreciate the challenge, consider this simple but  inaccurate way to select $U$. Given a desired latency ($r_0$) and the relation $r = f(u)$, which provides the steady-state latency $r$ of a server at an utilization $u$, choose $U = f^{-1}(r_0)$. The $U$ thus selected will likely result in a latency that is higher than the desired for several reasons such as imperfect load balancing, coordination overhead among peer caches and non-steady state behavior resulting from on-off transitions. The relation between $U$ and the latency of \shrink\ can be obtained with prior testing for a representative workload, thereby enabling operators to choose $U$ that maximizes energy savings while achieving desired latencies.

@@

%How does an operator select the utilization limit $U$ that maximizes energy savings while achieving the desired latency? A simple, but somewhat inaccurate, way to select $U$ is as follows. Given a desired latency ($r_0$) and a steady-state utilization-vs.-latency curve of a server $r = f(u)$, choose $U = f^{-1}(r_0)$. The utilization limit selected thus will likely result in a latency that is higher than the desired value due to imperfect load balancing and non-steady state behavior immediately after servers have been turned on or off. However, this utilization limit is expected to be a good starting point, after which operators can reduce $U$ in small decrements until latency reaches the desired value.

@@

Let $F(u)$ be the utilization vs. end-user latency curve for the operator's latency metric of interest provided to \shrink. $F(u)$ is similar to the end-to-end latency $(f(u) + mB + A)$ described in Section \ref{sec:analysis}. We discuss how $F(u)$ can be obtained in Section \ref{sec:utilization-vs-responsetime}.  \shrink\ uses $F(u)$ assuming that if the servers active in the datacenter are run at an average utilization $u$, then the resulting latency is $F(u)$. Thus, to achieve a specified end-user latency $R$, \shrink\ chooses a server utilization limit $U = r^{-1}(R)$.

%\shrink\ chooses a server utilization limit $U$ to achieve a specified end-user latency $R$.  If $F(u)$ is the utilization vs. end-user latency curve provided, $U = r^{-1}(R)$. $F(u)$ is similar to the end-to-end latency $(f(u) + mB + A)$ described in Section \ref{sec:analysis}. We discuss how $F(u)$ can be obtained in Section \ref{sec:utilization-vs-responsetime}.

@@

\subsection{Computing utilization vs. latency curve}
\label{sec:utilization-vs-responsetime}

We discuss three ways of computing the utilization vs. latency curve $F(u)$ input to \shrink\ in the order of increasing complexity. The more complex schemes potentially enable \shrink\ to provide latencies that are closer to those specified by operators.

The first and second approaches require prior measurements of a single server and of the entire \cdc\ respectively.
These measurements obtain the  value of the operator's latency metric of interest at varying levels of utilization.  These measurements are done with a representative workload in a test environment with real or emulated client-to-server delay and server-to-origin delay.  The first approach implicitly assumes that if the latency of a single server at utilization $u$ is $r$, then the latency of the \cdc\ whose servers have an average utilization $u$ is also $r$. This assumption ignores the inefficiencies of a distributed system such as imperfect load balancing and coordination overhead among peer caches in a \cdc. These factors are more accurately accounted for in the second approach. Nonetheless, both approaches could be inaccurate as they do not consider the non-steady cache behavior due to on-off transitions and increased cache miss rates due to reduced storage.  To account for these inaccuracies, a constant inefficiency factor $\rho$ is added as follows. If $F'(u)$ is the measured  utilization vs. latency curve, then \shrink\ is input a curve $F(u) = \rho \times F'(u)$.

The third approach is to equip \shrink\ to learn the curve automatically based on online measurements. The overhead of these measurements can be kept small by doing them for a small fraction of requests only. This approach could be more accurate than the previous two approaches because it considers the factors discussed above and can even adapt to changes in workload characteristics that influence the utilization vs. latency behavior. We have not implemented this third approach in \shrink.

%We discuss three ways of computing the utilization vs. latency curve that is provided to \shrink\ in the order of increasing complexity. The more complex schemes potentially enable \shrink\ to provide latencies that are closer to those specified by operators. The first and second approaches require prior measurements at varying levels of utilization of a single server and of the entire \cdc\  respectively. Similar to to the assumption made in deriving the model in Section \ref{sec:analysis}, the first approach implicitly assumes that the utilization vs. latency behavior of the datacenter is the same as that of a single server. However, a \cdc\ will likely have a higher latency than a single server at the same utilization because imperfect load balancing and coordination overhead among peer caches in a \cdc.  These factors are accounted for in the second approach; therefore, its utilization vs. latency curve is closer to the actual curve than the first approach.

%The first, and the simplest, approach is to compute the utilization vs. latency curve based on profiling a single server at varying levels of utilization. This approach implicitly assumes that the utilization vs. latency behavior of the datacenter is the same as that of a single server; we make the same assumption in Section \ref{sec:multiserver}.
